---
title: "Philosophy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{philosophy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library("projthis")
```

This package provides a set of tools to manage reproducible workflows that use RMarkdown.
The goal is to provide the "simplest way that works". 
That said, this package has three main ideas:

1. Package dependencies are managed using a DESCRIPTION file.

1. In this context, a "workflow" is a sequence of RMarkdown files. 
   To render an entire workflow is to render its RMarkdown files sequentially.

1. Workflow-rendering can be automated using GitHub Actions.

The first two ideas are independent of each other; the third idea builds on the first two.

There are more-comprehensive tools that address each of the first two ideas, for example: [renv](https://rstudio.github.io/renv/index.html) for package dependencies, and [targets](https://docs.ropensci.org/targets/) for workflow dependencies.
This package does *not* aspire to compete with these approaches. 
Rather, it aspires to be a first step folks can take towards these more-comprehensive approaches, as their needs may dictate.

This article takes a top-down approach to describe the philosophy of this package; other articles in this documentation aim to take a bottom-up approach.
This article assumes you are already familiar with the problems it is trying to solve, and some of the existing solutions.
 
## Main ideas

### DESCRIPTION file

For package development and deployment, the DESCRIPTION file is used to manage dependencies. 
Part of the CRAN "contract" dictates that your "stuff" work as you expect with the current CRAN versions of your dependencies (I realize I am being a little loose with terminology, but you get the idea).

This idea is extended to workflows: to declare all the dependencies in a DESCRIPTION file. 
This requires you to be open to the possibility (probability?) that you may have to amend your workflow in response to breaking changes in packages. 
If this is intended to be a "living" workflow, perhaps this is not too much of an imposition.

Further, this package uses `remotes::install_deps()` to install dependencies. 
The `Remotes` field can be used to specify GitHub and private repositories; this package does not have the means to populate the `Rempotes` field, so the analysis developer will have to do this manually.
Of course, it also remains the responsibility of the analysis developer to ensure that the analysis stays current with these package dependencies.

The purpose here is to establish a simple way to manage a project's package-dependencies. 
As your project advances, you may need a more sophisticated approach; [renv](https://rstudio.github.io/) gives you much more control to specify fixed versions of packages to use.

#### Implementation

- `proj_create()`: creates a new RStudio project with a DESCRIPTION file.
- `proj_use_description()`: creates a DESCRIPTION file, updates dependencies.
- `proj_check_deps()`, `proj_update_deps()`: checks (updates) dependencies listed in the DESCRIPTION file against those used in the project.
- `proj_install_deps()`: Install the package-dependncies listed in the DESCRIPTION file.

### Sequence of RMarkdown files

This package proposes a concept of a *workflow*, an overloaded word to be sure, but useful here. 
Here, a workflow consists of a directory containing a sequence of RMarkdown files and its own data directory.
An example:

```
important-workflow/
  data/
    00-import/
    01-do-stuff/
    02-do-more-stuff/
    99-publish/
  00-import.Rmd
  01-do-stuff.Rmd
  02-do-more-stuff.Rmd
  99-publish.Rmd
  README.Rmd
```

There are some rules governing the use of subdirectories of `data/`:

- Each RMarkdown file can write data to only one subdirectory: its *target* directory.

- An RMarkdown file can use as *source* directories only subdirectories created earlier in the workflow.

There are two special files in the workflow: `00-import.Rmd` and `99-publish.Rmd`:

- Use `00-import.Rmd` to import data into the workflow from the "rest of the universe". 

- Use `99-publish.Rmd` to manage the data that you wish to make available outside your workflow.

Using externally-accessible sites (e.g. Google Drive) from which to import data would make your workflow portable. 
To run your workflow on other computers, you would have to ensure access (credentials, etc.) to the external sites.

Similarly, you may wish to use the `99-publish.Rmd` file to post your "finished" data to an external site.

#### Implementation

There's a set of functions to manage workflows:

- `proj_use_workflow()`: creates a directory for a new workflow, including a `data` directory and a `README.Rmd` file (from a template).
- `proj_workflow_use_rmd()`: creates a new RMarkdown file (from a template) in your workflow.
- `proj_workflow_render()`: renders all the RMarkdown files in sequence, with `README.Rmd` last.

Do I 

The function `proj_workflow_use_rmd()` works best if called when the active file in your RStudio IDE is another RMarkdown file from the same workflow; this provides the directory in which to create the new file.

There's another set of functions to manage directories from within an RMarkdown file. 

- `proj_create_dir_target()`: create the target directory.
- `proj_path_source()`: function factory, creates function to access source paths.
- `proj_path_target()`: function factory, creates function to access target paths.
- `proj_dir_info()`: get concise version of data returned from `fs::dir_info()`.

The function `proj_create_dir_target()` has an argument `clean` which defaults to `TRUE`. 
Its purpose is to indicate if an existing target directory is to be deleted before it is created. 

There is one situation where I think it useful to specify `clean` as `FALSE`.
Let's say that the external source used by your `00-import.Rmd` file is a collection of daily files, with a new file added every day. 
You may not wish to download all the files every time the workflow is rendered. If your method for downloading files takes this into account, you may wish to 

The functions `proj_path_source()` and `proj_path_target()` are function factories; they return functions.
Within an RMarkdown file, you would use these returned functions to access paths in your source directories and your target directory.

For example:

```r
# path_source and path_target will be functions
path_source <- proj_path_source("01-do-stuff")
path_target <- proj_path_target("01-do-stuff")
```
These function factories each take a single argument, `name`, which shall correspond to the basename of the RMarkdown file that uses it.
The boilerplate code to support this is included in the template used with `proj_workflow_use_rmd()`.

In the rest of your RMarkdown file, you can use `path_source()` and `path_target()` in a similar way as you use `here::here()`: they provide you with locally-valid fully-qualified paths to the workflow's `data/` directory.

`path_source()` validates that the path requested is **earlier** in the workflow than the `name` used in the function factory; `path_target()` uses the `name` to help form the path:

```r
# returns: "/path/to/this/workflow/00-import/raw-data.csv"
path_source("00-import", "raw-data.csv") 

# throws error because this path is not earlier than "01-do-stuff"
path_source("02-do-more-stuff", "tidy-data.csv") 

# returns: "/path/to/this/workflow/01-do-stuff/clean-data.csv"
path_target("clean-data.csv") 
```



### GitHub Actions

The idea with Github Actions is to automate your workflow. When running your workflow

#### Implementation

- `proj_workflow_use_action()`:

## Supporting ideas
