---
title: "projthis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{projthis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library("projthis")
```

The goal of the projthis package is to help you establish project-based workflows. 
This involves: 

- Managing your project's package-dependencies.

- Managing the dependencies among files in your workflow.

- Automating the rendering of your workflow using GitHub Actions.
  
## Managing package-dependencies

One of the challenges of making a project portable *and* reproducible is keeping the dependencies current.
There are some comprehensive approaches to this problem; the best known are the [packrat](https://rstudio.github.io/packrat/) package, and its successor, [renv](https://rstudio.github.io/renv/).

The central idea of these packages is they assemble a set of packages that are particular to a given project, within that project. 
They build a package library dedicated to the project itself, offering a strong guarantee that the "same code" is run everywhere.

The central idea here is to offer a weaker guarantee, essentially the guarantee offered when submitting a package to CRAN: that the code will run using the latest versions of the packages listed in the `DESCRIPTION` file. 
It becomes the responsibility of everyone who contributes to the project to make sure that it runs as expected, and to make updates to the project should some subset of its dependencies behave differently. Again, this a variation on the CRAN "contract". 

This is a much-more lightweight approach as it requires only that the `DESCRIPTION` file be kept current.
If you have dependencies that are not on CRAN, you can use the `Remotes:` field, just as you would for a package repository.

### Workflow example

```r
proj_create()
```

```r
proj_use_description()
```

```r
proj_check_deps()
proj_update_deps()
```

```r
proj_install_deps()
```

Graduate to {renv}.

## Naming convention

Support of a workflow.

```r
# sets up directories, seeds files
proj_use_workflow()
# open _site.yml, index.Rmd
```

```r
# creates a new rmd file from a template
proj_workflow_use_rmd()
```

```r
# wraps rmarkdown::build_site()
proj_workflow_run()
```

```r
# adds a GitHub action to the repository
proj_workflow_use_action()
```

Graduate to {drake}, {targets}.

## Converting to package

You know the rules: copy-paste code three times, write a function. 
Copy-paste a function three times, write a package.

Use a legal name for the project. 
This leaves the door open to building a package on the same structure. 

Just like any other package, you would put the functions in the `R` directory, document, test, etc. 

Graduate to [R packages](https://r-pkgs.org/)

At some point, it may make sense to migrate the package away from the rest of the project, but that will be an exercise left to the reader.

## Converting to RMarkdown site

Who knows, maybe your project also demands a fancier, more structured website. 
Let's keep a path open for that, too.

I imagine that one repository, at some point, becomes two, and it will be up to the reader to manage that. I think the goal here is to set things up so that's as painless as possible.

Graduate to [rmarkdown site](https://bookdown.org/yihui/rmarkdown/rmarkdown-site.html), then on to Hugo, bookdown, distill, etc.
